---
title: "R Notebook"
output: html_notebook
---

# Data Frames

Matricies are like vectors-- they can only have one type of data in them. What if we want to handle more complex data?

#### Avengers Tenure
Let's revisit the Avengers data from PSet 1.
This is a small subset of a dataset compiled by Five Thirty Eight on deaths in the Marvel universe's Avengers. You can access the full dataset [here](https://data.world/fivethirtyeight/avengers) and the article it was used for [here](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/).


| Alias            | Appearances | Year joined | Full member? | Died | Returned | 
|------------------|-------------|-------------|--------------|------|----------| 
| She-Hulk         | 55          | 1982        | No           | No   |  N/A     | 
| Agent Anti-Venom | 746         | 2012        | No           | No   |  N/A     | 
| The Wasp         | 1165        | 1963        | Yes          | Yes  |  Yes     | 
| Storm            | 1598        | 2012        | Yes          | No   |  N/A     | 
| The Hulk         | 2089        | 1963        | Yes          | Yes  |  Yes     | 
| Thor             | 2402        | 1963        | Yes          | Yes  |  Yes     | 
| Iron Man         | 3068        | 1963        | Yes          | Yes  |  Yes     | 
| Wolverine        | 3130        | 2005        | Yes          | Yes  |  No      | 
| Spider-Man       | 4333        | 1990        | Yes          | Yes  |  Yes     | 

```{r}
aliases <- c("She-Hulk", "Agent Anti-Venom", "The Wasp", "Storm", "The Hulk", "Thor", "Iron Man", "Wolverine", "Spider-Man")
appearances <- c(55, 746, 1165, 1598, 2089, 2402, 3068, 3130, 4333)
years <- c(1982,2012,1963,2012,1963,1963,1963,2005,1995)
full <- c("No", "No", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes")
died <- c("No", "No", "Yes", "No", "Yes", "Yes", "Yes", "Yes", "Yes")
returned <- c("N/A", "N/A", "Yes", "N/A", "Yes", "Yes", "Yes", "No", "Yes")
```


```{r}
avengers <- data.frame("aliases" = aliases, "number of appearances" = appearances, "year joined" = years, "full member" = full, "died" = died, "returned" = returned)
avengers 
```

Stats on our new dataframe...
```{r}
dim(avengers)
colnames(avengers)
rownames(avengers)
```

Adding ID column
```{r}
avengers$id <- 1:nrow(avengers)
avengers
```


Adding tenure length column
```{r}
avengers$tenure <- 2020-avengers$year_joined
avengers
```


Making logical columns better
```{r}


```

Making still alive column

```{r}
avengers$still_alive <- !avengers$died | avengers$returned
avengers
```



Reordering columns
```{r}
col_order <- c("id", "aliases", "number of appearances", "year joined", "full member", "died")
avengers <- avengers[, col_order]
avengers
```


Querying
```{r}
avengers$id
avengers[8, ]
```







