---
title: "Lecture Code 2/12 - Dataframes, I/O, and Advanced Subsetting"
output: html_notebook
author: "Liz Fischer"
---

# Data Frames

Matricies are like vectors-- they can only have one type of data in them. What if we want to handle more complex data?

#### Avengers Tenure
Let's revisit the Avengers data from PSet 1.
This is a small subset of a dataset compiled by Five Thirty Eight on deaths in the Marvel universe's Avengers. You can access the full dataset [here](https://data.world/fivethirtyeight/avengers) and the article it was used for [here](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/).


| Alias            | Appearances | Year joined | Full member? | Died | Returned | 
|------------------|-------------|-------------|--------------|------|----------| 
| She-Hulk         | 55          | 1982        | No           | No   |  N/A     | 
| Agent Anti-Venom | 746         | 2012        | No           | No   |  N/A     | 
| The Wasp         | 1165        | 1963        | Yes          | Yes  |  Yes     | 
| Storm            | 1598        | 2012        | Yes          | No   |  N/A     | 
| The Hulk         | 2089        | 1963        | Yes          | Yes  |  Yes     | 
| Thor             | 2402        | 1963        | Yes          | Yes  |  Yes     | 
| Iron Man         | 3068        | 1963        | Yes          | Yes  |  Yes     | 
| Wolverine        | 3130        | 2005        | Yes          | Yes  |  No      | 
| Spider-Man       | 4333        | 1990        | Yes          | Yes  |  Yes     | 


Make the columns as vectors
```{r}
aliases <- c("She-Hulk", "Agent Anti-Venom", "The Wasp", "Storm", "The Hulk", "Thor", "Iron Man", "Wolverine", "Spider-Man")
appearances <- c(55, 746, 1165, 1598, 2089, 2402, 3068, 3130, 4333)
years <- c(1982,2012,1963,2012,1963,1963,1963,2005,1995)
full <- c("No", "No", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes")
died <- c("No", "No", "Yes", "No", "Yes", "Yes", "Yes", "Yes", "Yes")
returned <- c("N/A", "N/A", "Yes", "N/A", "Yes", "Yes", "Yes", "No", "Yes")
```

Create a data frame from the columns
```{r}
avengers <- data.frame("aliases" = aliases, "number of appearances" = appearances, "year joined" = years, "full member" = full, "died" = died, "returned" = returned, stringsAsFactors = FALSE)
avengers 
```

Stats on our new dataframe...
```{r}
dim(avengers)
colnames(avengers)
rownames(avengers)
```


Querying
```{r}
avengers$died #select a column
avengers[8, ] #select a row
avengers[8, "died"]
```

Replacing values
```{r}
# this is why stringsAsFactors = FALSE is important!

avengers[8, "died"] <- "banana"
```


Adding ID column
```{r}
avengers$id <- 1:nrow(avengers)
avengers
```


Reordering columns
```{r}
col_order <- c("id", "aliases", "number.of.appearances", "year.joined", "full.member", "died", "returned")
avengers <- avengers[, col_order]
avengers
```


Adding tenure length column
```{r}
avengers$tenure <- 2020-avengers$year.joined
avengers
```


Making logical columns better
```{r}
for (row in 1:nrow(avengers)){
  full_member <- avengers[row, "full.member"]
  if (full_member == "No"){
    avengers[row, "full.member"] <- "FALSE"
  } else {
    avengers[row, "full.member"] <- "TRUE"
  }
}
# avengers$full_member <- ifelse(avengers$full_member == "Yes", TRUE, FALSE)
avengers$died <- ifelse(avengers$died == "No", FALSE, TRUE)
avengers$returned <- ifelse(avengers$returned == "Yes", TRUE, FALSE)
avengers
```

Making still alive column
```{r}
avengers$still.alive <- !avengers$died | avengers$returned
avengers
```


Reordering columns & omitting some
```{r}
col_order <- c("id", "aliases", "number.of.appearances", "tenure", "full.member", "still.alive")
avengers <- avengers[, col_order]
avengers
```



# I/O - Reading and Writing Files

## Reading
Finding our place on the computer
```{r}
getwd()
#setwd("/Desktop/Spring-2020-Lecture-Code")
#getwd()
```

Seeing what's there
```{r}
dir()
```


```{r}
playlist <- read.csv(file="spotify.csv", as.is=TRUE) #as.is=TRUE is another way to stop R from creating factors, like stringsAsFactors = FALSE.
playlist
```

## Writing
```{r}
write.csv(x=avengers, file="avengers.csv")
```


## Advanced subsetting

We can search a column for a value....
```{r}
playlist$artists == "ABBA"
```


Or multiple values...
```{r}
index <- (playlist$artists == "ABBA" | playlist$artists == "Cher")
index

#playlist[index,]
```


```{r}
playlist$artists %in% c("ABBA", "Cher", "Shakira")

#playlist[]
```

What happened to Shakira?
```{r}
playlist[grep("Shakira", playlist$artists),]

#playlist[grep("Shakira|ABBA|Cher", playlist$artists),] # this is taking advantage of Regular Expressions, which you are free to read about as needed but we won't cover extensively


```

Self-titled albums
```{r}
playlist[playlist$album %in% playlist$artists, ]
```

Reorder
```{r}
index <- order(playlist$popularity, decreasing=TRUE)
index

#playlist[index,]
```




We'll talk about combining tables later, but read the book section that talks about it (2.9)

