---
title: "Code from class 2/17"
output: html_notebook
---

Let's load some data
```{r}
books <-  read.csv("goodreads.csv", as.is=TRUE)
books
```

Let's look at language. How what's the distribution like?
```{r paged.print=TRUE}
tab <- table(books$language_code)
tab
```

We can reorder these to make it a bit easier to see...
```{r}
tab <- tab[order(tab, decreasing = TRUE)]
tab
```

Or get just the counts...
```{r}
counts <- as.numeric(tab)
class(counts)
```

Or show just the lables, ordered from fewest to most occurances 
```{r}
names(tab)[order(tab,decreasing = FALSE)]
```

Let's combine all the English variations into one.
```{r}
books$language_code <- ifelse(books$language_code == "en-US" | books$language_code == "en-GB" | books$language_code == "en-CA", "eng", books$language_code)

#OR books$language_code <- ifelse(grepl("en-", books$language_code), "eng", books$language_code)

books$language_code <- ifelse(grepl("en-", books$language_code),"eng",
books$language_code)
books
```




Let's look at length. What is the distribution like?
```{r}
hist(books$X..num_pages)
```

Whoa, that's a wide span. Hard to tell much detail like that.... Let's try to eliminate some outliers

```{r}
?hist
hist(books$X..num_pages[books$X..num_pages <2000], breaks = 50, xlab = "Number of Pages", main = "Book length (under 2000 pages)")

```

```{r}
quantile(books$X..num_pages)
```


