---
title: "2/24: Binning, Combining Plots, Aggregation, Applying Functions"
output: html_notebook
---

```{r}
df <- read.csv("ks-projects-201801.csv", as.is=TRUE)
t <-table(df$category)
t[order(t)]
```

Let's get just Video Games for a moment

```{r}
vg <- df[df$category == "Video Games", ]
table(vg$state)[order(table(vg$state))]
```

Adding a year column to make the data easier to work with
```{r}
for (i in 1:length(vg$launched)) {
  x <- strsplit(vg$launched[i], "-")
  vg$year[i] <- as.numeric(x[[1]][1])
}
```


#Histograms

See number of video games each year
```{r}
vg$year <- as.numeric(vg$year)
table(vg$year)
hist(vg$year)
```

Calcualting & visualizing average pledge size
```{r}
avgPledge <- vg$usd_pledged_real/vg$backers
avgPledge[is.nan(avgPledge) | is.infinite(avgPledge)] <- 0

hist(avgPledge[avgPledge <200], breaks=50, col="gray")
```



# QUANTILES

```{r}
quantile(vg$usd_goal_real)
```

Notice the positive skew


```{r}
coff <- quantile(avgPledge, prob=.25)
coff
lowPledgeFlag <- (avgPledge < coff)
table(lowPledgeFlag, vg$state)
```


# Binning
```{r}
breakPoints <- quantile(avgPledge, prob=seq(0,1,length.out=11), names=FALSE)
breakPoints
```

Putting each row into one of the bins by quantile
```{r}
bin <- cut(avgPledge, breakPoints, labels=FALSE, include.lowest = TRUE)
table(bin, vg$state)
```

Equal-area histogram using quantiles:
```{r}
breakPoints[1:10]
hist(avgPledge[avgPledge < 80], breaks=breakPoints[1:10])
```

```{r}
vg$avg <- vg$usd_pledged_real/vg$backers
```

# Combining Plots

```{r}
t <- table(vg$year)
years <-as.numeric(names(t))

par(mfrow=c(3,3))
#par(mar=c(0,0,2,0))

for (i in 1:length(years)){
  y <- vg[vg$year == years[i], ]
  #print(y)
  #print(y$avg[y$avg < 100])
  hist(y$avg[y$avg < 100], breaks=50, ylim = c(0, 90), col = "dodgerblue", main=paste("Average pledges in", years[i]), xlab="", ylab="", axes = FALSE)
  box()
}

```



#Aggregation

```{r}
s <- df[df$state=="successful",]

categories <- unique(s$category)

moneySpent <- 1:length(categories)
names(moneySpent) <- categories

for (c in categories){
  index <- which(df$category == c)
  moneySpent[c] <- sum(df[index,"usd_pledged_real"])
}


moneySpent[order(moneySpent, decreasing = TRUE)]

```

# Applying functions

```{r}
apply(df[,15:14], 2, sum)

```

```{r}
debates <- read.csv("debate_transcripts.csv", as.is=TRUE)
```

```{r}
speakers <- unique(debates$speaker)
speakers
```

```{r}
x <- tapply(debates$speech, INDEX=debates$speaker, FUN=paste, collapse=" ")

library(stringr)
str_count(x["George S."], "Sanders")

apply(x, 1, FUN=str_count, pattern="Sanders|Bernie")

```





