---
title: "3/4: Better Scatter Plots, Correlation, and Linear Models"
output: html_notebook
---

#Prepping data
Load the data
```{r}
df <- read.csv("ks-projects-201801.csv", as.is=TRUE)

```

Adding useful columns
```{r}
df$average_pledge <- df$usd_pledged_real/df$backers
df$average_pledge <- ifelse(is.nan(df$average_pledge), 0, df$average_pledge)

df$percent_of_goal <- round(df$usd_pledged_real/df$usd_goal_real, digits=3)
```


#Scatter plots
```{r}
art <- df[df$main_category=="Art" & df$state=="successful",]

mid <- art[
  art$usd_goal_real > quantile(art$usd_goal_real, prob=.1) &
  art$usd_goal_real < quantile(art$usd_goal_real, prob=.9) &
  art$usd_pledged_real > quantile(art$usd_pledged_real, prob=.1) &
  art$usd_pledged_real < quantile(art$usd_pledged_real, prob=.9)
,]

plot(mid$usd_goal_real, mid$usd_pledged_real)
```

```{r}
pchVals <- rep(1, nrow(mid))
pchVals[mid$category=="Public Art"] <- 3

plot(mid$usd_goal_real, mid$usd_pledged_real, pch=pchVals)
```

```{r}
colorVals <-  rep("black", nrow(mid))
colorVals[mid$category=="Public Art"] <- "red"

plot(mid$usd_goal_real, mid$usd_pledged_real, pch=pchVals, col=colorVals)
```


```{r}
lowPledge <- quantile(mid$average_pledge, prob=.1)

colorVals <- rep("black", nrow(mid))
colorVals[mid$average_pledge <= lowPledge] <- "green"

plot(mid$usd_goal_real, mid$usd_pledged_real, col=colorVals)
```


Let's say we want color-code the subcategories of Art...
```{r}
subcats <- unique(art$category)
```

First we make these numbers...
```{r}
categoryIndex <- match(art$category, subcats)
categoryIndex
```
Then we assign colors to each number...
```{r}
colors <-  c("green", "red", "blue", "orange", "purple", "black", "brown", "aquamarine1", "coral", "grey", "firebrick", "beige", "cornsilk")
#Or get random colors with: 
#colors <-  sample(colors(), length(subcats))
#colors

colors[categoryIndex]

```
And use them in a plot
```{r}
plot(mid$usd_goal_real, mid$usd_pledged_real, col=colors[categoryIndex], pch=19)
```

#Adding text
Let's say we want to do something with the subcategories of Art...
```{r}
unique(art$category)
```

Get average amount pledged in each subcategory
```{r}
catPledgedAvg <- tapply(art$usd_pledged_real, art$category, mean)
catPledgedAvg
```

And average goal in each subcategory
```{r}
catGoalAvg <- tapply(art$usd_goal_real, art$category, mean)
catGoalAvg
```

Then plot those
```{r}
plot(catGoalAvg, catPledgedAvg)
```

Can add labels
```{r}
plot(catGoalAvg, catPledgedAvg)
text(catGoalAvg, catPledgedAvg-300, labels=names(catGoalAvg))
```

# Correlation
The closer it is to +/- 1, the stronger the relationship
```{r}
df[, c("usd_goal_real", "usd_pledged_real")]
#cor(df[, c("usd_goal_real", "usd_pledged_real")])
#cor(mid[, c("usd_goal_real", "usd_pledged_real")])
```

#Best fit line
```{r}
lm(mid$usd_pledged_real ~ mid$usd_goal_real)
```
```{r}
plot(mid$usd_goal_real, mid$usd_pledged_real)
abline(593.970, 1.053)
```
